---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(ggplot2)
data<-read.csv("activity.csv")
stepsdata <- na.omit(data) #remove NA data
stepsdata$date <- as.Date(stepsdata$date) #convert date into factor
```



## What is mean total number of steps taken per day?
###calculate the total number of steps taken per day

```{r}
y<-aggregate(x = stepsdata$steps,                # Specify data column
          by = list(stepsdata$date),              # Specify group indicator
          FUN = sum) 
head(y,10)
``` 
###histogram for the total number of steps per day

```{r}
hist(y$x, 
     xlab="Total number of steps taken each day", 
     ylab="Count", 
     main="Histogram of total number of steps taken each day",
     col=5)
```
###Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(y$x,na.rm=TRUE)
median(y$x,na.rm=TRUE)
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
