---
title: "Reproducible Research: Peer Assessment 1 by JANVI SAHU"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
library(ggplot2)
data<-unzip("activity.zip")
data<-read.csv("activity.csv")
stepsdata <- na.omit(data) #remove NA data
stepsdata$date <- as.Date(stepsdata$date) #convert date into factor
```

## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day

```{r}
y<-aggregate(x = stepsdata$steps,                # Specify data column
          by = list(stepsdata$date),              # Specify group indicator
          FUN = sum) 
head(y,10)
```

Histogram for the total number of steps per day

```{r}
hist(y$x, 
     xlab="Total number of steps taken each day", 
     ylab="Count", 
     main="Histogram of total number of steps taken each day",
     col=5)
```

\#\#\#Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(y$x,na.rm=TRUE)
median(y$x,na.rm=TRUE)
```

## What is the average daily activity pattern?

Find mean steps taken grouped by interval

```{r}
library(dplyr)
z<-stepsdata %>%
group_by(interval) %>%
summarize(mean_steps= mean(steps))
z

```

Time series plot

```{r}
ggplot(z, aes(x = interval , y = mean_steps)) + geom_line(color="sky blue", size=1) + labs(title = "Avg. Daily Steps", x = "Interval", y = "Avg. Steps per day")
```

Find the interval with maximum mean of steps

```{r}

maxstepinterval <- z$interval[which.max(z$mean_steps)]
maxstepinterval
```

## Imputing missing values
The number of missing values

```{r}
a<-nrow(data)
b<-nrow(stepsdata)
a-b
```

The number of missing values 2304

 Create a filled in dataset by assigning the average value for that time interval is NA is found.
```{r}
filldata <-data
for (i in 1:nrow(filldata)) {
    if (is.na(filldata$steps[i])) {
        # Find the index value for when the interval matches the average
        index <- which(filldata$interval[i] ==z$interval)
        # Assign the value to replace the NA
        filldata$steps[i] <- z[index,]$mean_steps
    }
}
filldata$date <- as.Date(filldata$date) 
```
Histogram of the total number of steps taken each day
```{r}
h<-filldata %>%
group_by(date) %>%
summarize(total_steps= sum(steps))

hist( h$total_steps, xlab="Number of steps", ylab="count", main="Histogram of number of steps")
```
MEAN AND MEDIAN OF STEPS IN THE NEW DATASET
```{r}
fillmeansteps <- mean(h$total_steps)
fillmediansteps <- median(h$total_steps
                          )
fillmediansteps

```

The mean total number of steps per day is 10766.19 and the median is 10766.19.
## Are there differences in activity patterns between weekdays and weekends?
```{r}



# Make weekday variable
filldata$day <- weekdays(filldata$date)
# Define all days as weekdays
filldata$daytype <- "weekday"
# Fix days that are saturday or sunday to be weekends
filldata$daytype[filldata$day %in% c("Saturday", "Sunday")] <- "weekend"
```
Calculate the average weekday steps versus average weekend steps
```{r}



# Group data by 5 minute interval and summarize the average
# number of steps in that interval
dayaverage <- filldata %>%
    group_by(daytype, interval) %>%
    summarize(AverageSteps=mean(steps))
```

    
```{r}
qplot(interval, AverageSteps, data=dayaverage,
      type="l",
      geom="line",
      xlab="Interval",
      ylab="Number of Steps (Average)",
      main="Average steps taken Weekends vs. Weekdays",
      facets =daytype ~ .)
```


